---
title: "Just One Guy's Introduction to Shiny"
author: "Jay Emerson"
date: "Jaks Tap, Chicago, November 13, 2014"
output: ioslides_presentation
runtime: shiny
---


## This Presentation

- Official resources
- Toy example (stand-alone)
- Toy example (via `ShinyHelper`) embedded in this presentation (method 1)
- Slightly expanded example embedded in this presentation (method 2)


## Where To Go (official resources)

Googling can reveal some older tutorials and documentation.
None of this is my work;  my stuff is of secondary importance
in the grand scheme of things and will be referenced later, separately.
The newest and greatest official resources, as far as I know:

- Tutorial, documentation, examples: http://shiny.rstudio.com
- Source code (and see `inst/examples/` for
  some good example): https://github.com/rstudio/shiny
- Serious deployment (Alpha version): https://www.shinyapps.io


## 01_hello/ folder contents

This example is included with the `shiny` package in the `inst/examples`
directory along with other goodies.  We'll look at it, then we'll run it.

    Jays-MacBook-Pro:Shiny_master jay$ ls 01_hello/
    
    server.R  ui.R

The following code is mildly edited to display well in the presentation.


## 01_hello/ui.R
```{r eval=FALSE}
library(shiny)

shinyUI(fluidPage(
  titlePanel("Hello Shiny!"),
  sidebarLayout(
    sidebarPanel(
      sliderInput("bins", "Number of bins:",
                  min = 1, max = 50, value = 30)
    ),
    mainPanel(
      plotOutput("distPlot")
    )
  )
))
```
    
    
## 01_hello/server.R
```{r eval=FALSE}
library(shiny)

# Define server logic required to draw a histogram
shinyServer(function(input, output) {
  
  output$distPlot <- renderPlot({
    x    <- faithful[, 2]  # Old Faithful Geyser data
    bins <- seq(min(x), max(x), length.out = input$bins + 1)
    
    hist(x, breaks = bins, col = 'darkgray', border = 'white')
  })
  
})
```


## Running 01_hello in browser
    Jays-MacBook-Pro:Shiny_master jay$ ls
    
    01_hello  Intro2Shiny.Rmd
    
    Jays-MacBook-Pro:Shiny_master jay$ ls 01_hello/
    
    server.R  ui.R
    
    Jays-MacBook-Pro:Shiny_master jay$ R

      < snip -- I'm still apparently using R 3.1.1 >

    > library(shiny)
    > runApp("01_hello/")

    Listening on http://127.0.0.1:7934
    ^C
    >  

```{r echo=FALSE}
unlink("SH_hello/", recursive=TRUE)
```

## ShinyHelper:

__Author simple Shiny apps with wrapper functions provided
by `ShinyHelper`.__

If you just want to build an occasional
Shiny app, you probably don't need this.  But if it becomes
a regular thing, `ShinyHelper` lets you write an app using
R code (rather than writing the app itself).  Contributions
would be welcome (providing new wrappers for more advanced
Shiny features).

https://github.com/jayemerson/ShinyHelper

Buyer beware, no warranties, etc...


## 01_hello via ShinyHelper

```{r eval=FALSE}
require(devtools)
install_github('ShinyHelper', 'jayemerson')
```
```{r}

library(ShinyHelper)

shinySkeleton("SH_hello")
headerPanelAdd("01_hello via ShinyHelper")
sliderInputAdd("bins", "Number of bins:",
               min = 1, max = 50, value = 30)
plotOutputAdd("distplot", plotcode =
'
  x    <- faithful[, 2]  # Old Faithful Geyser data
  bins <- seq(min(x), max(x), length.out = input$bins + 1)
  hist(x, breaks = bins, col = "darkgray", border = "white")
')
createApp("SH_hello", overwrite=TRUE)
# runApp("SH_hello") # We'll do something else...
```


## Embedding SH_hello in this presentation

You can simply point your presentation to the folder
containing your Shiny app.  In this example,
the following code goes into a simple code chunk
on the next slide that looks like this (but with `echo=FALSE`
as a code chunk option):

```{r eval=FALSE}
shinyAppDir(
  "SH_hello", options=list(width="100%", height=700)
)
```    


## SH_hello

```{r, echo = FALSE}
shinyAppDir(
  "SH_hello", options=list(width="100%", height=700)
)
```


## Embedding the Shiny app code inline

The example on the next pages comes from the default
page template for a Shiny R Markdown presentation.
See also:

http://rmarkdown.rstudio.com/authoring_shiny.html

Note that this approach doesn't use (require?) the "packaging"
of `inputPanel()` and `renderPlot()` inside formal
`shinyServer()` and `shinyUI` infrastructure.  There
is no formal directory containing the app, as with previous
examples.


## Embedded code for interactive plot

```{r, eval=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE,
       breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)",
       main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Slide with Interactive Plot

```{r, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE,
       breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)",
       main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

